Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-09 ICML-submission/related-work.tex
Rename linear separability to weak linear separability

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\cite{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the standard
linearly separable setting are proposed: one with a mistake bound of $O(K^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret against the multiclass hinge loss
in the general setting, and has a $O(\sqrt{T})$ mistake
bound in the $\gamma$-weak linearly separable setting. The polynomial dependence
of \textsc{Banditron}'s mistake bound on the time horizon is undesirable for
problems with a long time horizon. One key open question left
by~\cite{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can design
computationally efficient algorithms that achieve mistake bounds that match or
improve over those of inefficient algorithms. In this paper, we take a step
towards answering this question, showing that efficient algorithms with mistake
bounds quasipolynomial in the margin parameter can be obtained.

@@

Many works consider bandit multiclass classification in the general
non-separable setting. \cite{Abernethy-Rakhlin-2009} poses the open problem of
whether one can design an efficient algorithm to get a $O(\sqrt{T})$ regret
against some reasonable loss functions. \cite{Crammer-Gentile-2013} show that
such an algorithm can be obtained, provided that the distribution of
the label $y_t$ conditioned on feature vector $x_t$ satisfies certain parametric
noise assumption. \cite{Hazan-Kale-2011} developed the \textsc{Newtron}
algorithm, which has a regret between $O(\ln T)$ and $O(T^{2/3})$ against the
multiclass logistic loss, where the exact order of the regret depends on the
diameter of the benchmark class. In particular, if the diameter is $O(\ln T)$,
its regret bound would become $O(T^{2/3})$. The \textsc{SOBA} algorithm by
\cite{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
$\widetilde{O}(\sqrt{T})$ against the
$\eta$-loss~\cite{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, its regret
bound does not depend sensitively on the diameter of the benchmark class.
\cite{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
it doubly-exponentially improves over \cite{Hazan-Kale-2011}'s regret on its
dependence on the diameter of the benchmark class.

\cite{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass
boosting under bandit feedback, where one can view online boosting as online
linear classification by treating each base hypothesis as a separate feature.
\cite{Chen-Lin-Lu-2014}'s online weak learning condition implies that
the set of examples is strongly separable by a convex combination of base hypotheses
with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
mistake bound. \cite{Zhang-Jung-Tewari-2018} considers an online weak learning
condition, which implies that all examples is separable by a convex combination of base hypotheses with a margin (See~\cite{Mukherjee-Schapire-2013}, Theorem 3).
Under this condition, it gives
a boosting algorithm with a $O(\sqrt{T})$ mistake bound with the knowledge of the edge
parameter. In addition, it gives an adaptive algorithm with a $O(T^{3/4})$
mistake bound.

The online bandit multiclass learning is a special case of the contextual bandits
problem~\cite{Auer-2003, Langford-Zhang-2008}, where the cost is the classification error.
A series of works in the contextual bandits learning literature focuses on
oracle-efficiency, that is, they assume access to an
oracle that given a set of cost-sensitve learning examples,

@@

%as algorithms generate cost-sensitive examples that may not be linearly separable
%Algorithms that satisfy oracle efficiency has been proposed in the literature, for
%example.

Recently, \cite{Foster-Krishnamurthy-2018} developed a rich theory of contextual bandits
with surrogate losses, focusing on benchmarks of the form $\min_{f \in \calF}
\sum_{t=1}^T \frac{1}{K} \sum_{i=1}^K c_t(i) \phi( f_i(x_t) )$,
where $\calF$ contains score functions $f$
such that $\sum_{i=1}^K f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s
\gamma, 0)$ or $\min(1, \max(1 - \frac s \gamma, 0))$. On one hand, it gives
information-theoretic mistake upper bounds under various settings of $\calF$,
including parametric and nonparametric classes. On the other hand, it gives
an efficient algorithm that has a $O(\sqrt{T})$ regret against the benchmark of
$\calF = \cbr{x \mapsto W x: W \in \R^{k \times d}, \one^T W = 0}$. A direct
application of this result to the online bandit multiclass learning problem
yields an algorithm with a $O(\sqrt{T})$ mistake bound in the strong
linearly separable setting.

>>>>

The problem of bandit multiclass linear classification was initially formulated
by the pioneering work of~\citet{Kakade-Shalev-Shwartz-Tewari-2008}. In that
work, two computationally inefficient algorithms that work in the standard
linearly separable setting are proposed: one with a mistake bound of $O(K^2 d
\ln(\frac{d}{\gamma^2}))$, the other with a mistake bound of
$\widetilde{O}(\frac{K^2}{\gamma^2} \ln T)$. In addition, the authors propose
the \textsc{Banditron} algorithm, a computationally efficient algorithm that has
a $O(T^{2/3})$ regret against the multiclass hinge loss in the general setting,
and has a $O(\sqrt{T})$ mistake bound in the $\gamma$-weak linearly separable
setting. The polynomial dependence of \textsc{Banditron}'s mistake bound on the
time horizon is undesirable for problems with a long time horizon. One key open
question left by~\citet{Kakade-Shalev-Shwartz-Tewari-2008} is whether one can
design computationally efficient algorithms that achieve mistake bounds that
match or improve over those of inefficient algorithms. In this paper, we take a
step towards answering this question, showing that efficient algorithms with
mistake bounds quasipolynomial in the margin parameter can be obtained.

@@

Many works consider bandit multiclass classification in the general
non-separable setting. \citet{Abernethy-Rakhlin-2009} poses the open problem of
whether one can design an efficient algorithm to get a $O(\sqrt{T})$ regret
against some reasonable loss functions. \citet{Crammer-Gentile-2013} show that
such an algorithm can be obtained, provided that the distribution of the label
$y_t$ conditioned on feature vector $x_t$ satisfies certain parametric noise
assumption. \citet{Hazan-Kale-2011} developed the \textsc{Newtron} algorithm,
which has a regret between $O(\ln T)$ and $O(T^{2/3})$ against the multiclass
logistic loss, where the exact order of the regret depends on the diameter of
the benchmark class. In particular, if the diameter is $O(\ln T)$, its regret
bound would become $O(T^{2/3})$. The \textsc{SOBA} algorithm by
\citet{Beygelzimer-Orabona-Zhang-2017} achieves a regret of
$\widetilde{O}(\sqrt{T})$ against the
$\eta$-loss~\citet{Orabona-Cesa-Bianchi-Gentile-2012}. In addition, its regret
bound does not depend sensitively on the diameter of the benchmark class.
\citet{Foster-Kale-Luo-Mohri-Sridharan-2018} developed an algorithm that has a
regret of $\widetilde{O}(\sqrt{T})$ against the multiclass logistic loss, where
it doubly-exponentially improves over \citet{Hazan-Kale-2011}'s regret on its
dependence on the diameter of the benchmark class.

\citet{Chen-Lin-Lu-2014, Zhang-Jung-Tewari-2018} study online bandit multiclass
boosting under bandit feedback, where one can view online boosting as online
linear classification by treating each base hypothesis as a separate feature.
\citet{Chen-Lin-Lu-2014}'s online weak learning condition implies that the set
of examples is strongly separable by a convex combination of base hypotheses
with a margin. Under this condition, it gives an algorithm with a $O(T^{3/4})$
mistake bound. \citet{Zhang-Jung-Tewari-2018} considers an online weak learning
condition, which implies that all examples is separable by a convex combination
of base hypotheses with a margin (See~\cite{Mukherjee-Schapire-2013}, Theorem
3). Under this condition, it gives a boosting algorithm with a $O(\sqrt{T})$
mistake bound with the knowledge of the edge parameter. In addition, it gives an
adaptive algorithm with a $O(T^{3/4})$ mistake bound.

The online bandit multiclass learning is a special case of the contextual
bandits problem~\citep{Auer-2003, Langford-Zhang-2008}, where the cost is the
classification error. A series of works in the contextual bandits learning
literature focuses on oracle-efficiency, that is, they assume access to an
oracle that given a set of cost-sensitve learning examples.

@@

%as algorithms generate cost-sensitive examples that may not be linearly
%separable Algorithms that satisfy oracle efficiency has been proposed in the
%literature, for example.

Recently, \citet{Foster-Krishnamurthy-2018} developed a rich theory of
contextual bandits with surrogate losses, focusing on benchmarks of the form
$\min_{f \in \calF} \sum_{t=1}^T \frac{1}{K} \sum_{i=1}^K c_t(i) \phi( f_i(x_t)
)$, where $\calF$ contains score functions $f$ such that $\sum_{i=1}^K
f_i(\cdot) \equiv 0$, and $\phi(s) = \max(1 - \frac s \gamma, 0)$ or $\min(1,
\max(1 - \frac s \gamma, 0))$. On one hand, it gives information-theoretic
mistake upper bounds under various settings of $\calF$, including parametric and
nonparametric classes. On the other hand, it gives an efficient algorithm that
has a $O(\sqrt{T})$ regret against the benchmark of $\calF = \cbr{x \mapsto W x:
W \in \R^{k \times d}, \one^T W = 0}$. A direct application of this result to
the online bandit multiclass learning problem yields an algorithm with a
$O(\sqrt{T})$ mistake bound in the strong linearly separable setting.

### 2019-01-09 arxiv-version/introduction.tex
Rename linear separability to weak linear separability

In this paper, we focus on the special case when the examples lie in $\R^d$ and
are \emph{linearly separable with a margin}.
\chicheng{I wonder if we should introduce the two notions of linear separability here,
because strong linear separability is also worth mentioning in the full-information setting.
}
\david{The formal definitions are in a section of their own.}

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Beside the standard notion of linear separability, we define also
\emph{strong linear separability}. For binary classification, these two
definitions are identical. However, they differ for multiclass classification.
For multiclass classification with $K$ classes the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane. We state the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

For the case when the examples are strongly linearly separable, we design a
simple and efficient algorithm with expected number of mistakes at most $(K-1)
\lfloor (R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies
of the \textsc{Binary Perceptron} algorithm, one copy for each class.
Intuitively speaking, the factor $K-1$ is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We also
prove that any (possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs. First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernelized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
and memory complexity.

These algorithms differ by choice of the mapping or, equivalently, choice of the
kernel. The crux of the problem is the analysis of the approximation properties
of these mappings/kernels and the way margin in the original space $\R^d$ under
standard linear separability notion gets transformed into a margin in the
high-dimensional space under the strong separability notion. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \cite{Klivans-Servedio-2008}. As an important technical
achievement, we improve on the results of \cite{Klivans-Servedio-2008} by
removing completely the dependency on the original dimension $d$.

>>>>

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are \emph{linearly separable with a margin} (see
\autoref{definition:linear-separability} for a formal description).
%\chicheng{I wonder if we should introduce the two notions of linear separability here,
%because strong linear separability is also worth mentioning in the full-information setting.
%}
%\david{The formal definitions are in a section of their own.}

@@

The bandit feedback setting, however, is much more challenging. To understand
the differences, we dissect the notion of linear separability in the multiclass
setting. Besides the standard notion of linear separability, we also consider
\emph{strong linear separability} (see
\autoref{definition:strong-linear-separability} for a formal description). For
binary classification, these two definitions are identical. However, they differ
for multiclass classification with $K \geq 3$. For multiclass classification
with $K$ classes, the standard notion of linear separability means that the
examples from each class lie in an intersection of $K-1$ halfspaces and the
examples outside of the class lie in the complement of the intersection of the
halfspaces. Strong linear separability means that examples from each class are
separated from the remaining examples by a \emph{single} hyperplane. We state
the definitions precisely in
Section~\ref{section:notions-of-linear-separability}.

For the case when the examples are strongly linearly separable, we design a
simple and efficient algorithm with an expected number of mistakes at most $(K-1)
\lfloor (R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies
of the \textsc{Binary Perceptron} algorithm, one copy for each class.
Intuitively speaking, the factor $K-1$ is the price we pay for the bandit
feedback, or more precisely, the lack of full-information feedback. We also
prove that any (possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$
mistakes in the worst case.

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs.
\chicheng{I wonder if we should just present the results using the rational kernel,
 as results using other kernels are nonadaptive,
 and achieve qualitatively similar mistake bounds.}
First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernelized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
and memory complexity.

These algorithms differ by choice of the mapping or, equivalently, choice of the
kernel. The crux of the problem is the analysis of the approximation properties
of these mappings/kernels and the way margin in the original space $\R^d$ under
standard linear separability notion gets transformed into a margin in the
high-dimensional space under the strong separability notion. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \citet{Klivans-Servedio-2008}. As an important technical
achievement, we improve on the results of \citet{Klivans-Servedio-2008} by
removing completely the dependency on the original dimension $d$.
